Inferring neural networks from fluorescent calcium imaging using partial correlation and sample weighting

### 2014-05-28 article.tex
wording

% BG : brain, complex organ
Understanding how the brain works is a key to understand and to treat
brain pathologies and disorders such as Parkinson's disease, epilepsy.
Retrieving the connectomes, neurons connectivity map, will shed new lights on
the anatomical and functional connectivity of the brain.
The human brain is a complex biological organ, which is formed by 100
billions of neurons with 7000 synaptic connections on average. Inferring the
neuron connectivity network through dissection would be an impossible
and a daunting task.

% Issue : we have data, but need to have to infer the connectome
Cutting edge optical devices [todo cite] allows to track simultaneously
the neural activity of thousands of neurons through the fluorescent
calcium indicator. Those devices have however a sampling rate smaller
than the neuron firing speed and takes only two dimensional image, which may lead
to superpose the activity of different neurons. Furthermore, the
decay of the fluorescent calcium signal is slow compare to its rise.
The neural connectivity inference consists in highlighting, from observational
data, direct relationships between neurons. One way to emphasize the
connections between neurons is to represent them by a (un)directed graph with
$p$ nodes, where each node represents a neuron, and an edge $y_{ij}$ which
represents a direct link from neuron  $i$ to neuron $j$. If the graph is
undirected, then $y_{ij}$ is equal to $y_{ji}$  for all $i,j$.

@@

@@

The challenge was stated as an unsupervised learning problem. The particularity
of unsupervised inference is the network recovery solely based on
observational data. Other approaches have been considered to infer biochemical
networks, e.g. gene regulatory network, such as supervised learning [todo cite]
or transfer learning [todo cite], but these learning processes require, as
starting point, a partial knowledge of the target network or known networks
similar to the unknown one.

>>>>

% BG : brain, complex organ
Understanding how the brain works is a key to understand and to treat
brain pathologies and disorders such as Parkinson's disease and epilepsy.
Retrieving the connectomes, neurons connectivity map, will shed new lights on
the anatomical and functional connectivity of the brain. However,
the human brain is a complex biological organ, which is formed by 100
billions of neurons with 7000 synaptic connections on average. Inferring the
neuron connectivity network through dissection would be an impossible
and a daunting task.

Cutting edge optical devices [todo cite] allows to track simultaneously
the neural activity of thousands of neurons through the fluorescent
calcium indicator.
From those observational data, the neural connectivity inference consists in
highlighting direct relationships between neurons. However, those imaging
systems have a sampling rate smaller than the neuron firing speed and are
limited only two dimensional image, which may lead to superpose the activity
of different neurons. Furthermore, the decay of the fluorescent calcium signal
is slow compare to its rise.

Since only observational data are available, the network recovery is set as an
unsupervised learning problem. Whenever a partial knowledge of the target
network or known networks highly similar to the unknown one is available,
supervised learning [todo cite] and transfer learning  [todo cite] based
approach could be considered to infer networks.

@@

One way to emphasize the connections between neurons is to represent them by a
(un)directed graph with $p$ nodes: each node represents a neuron, and each
edge $y_{ij}$ represents a direct link from neuron  $i$ to neuron $j$. If the
graph is undirected, then $y_{ij}$ is equal to $y_{ji}$  for all $i,j$.

@@

From a probabilistic point of view, the newtwork

% Plan:
% introduce the notion of dependance/independance
% Introduce correlation => Estimator of I(X; Y)
%                       => Issue indirect effect
% Introduce partial correlation => Estimator of I(X; Y | Z)
%                               => Symmetric measure
% Introduce a custom solution to make the Y matrix assymetric
% Introduce Filtering
